Plant-pollintor networks worldwide are composed by the same specific building blocks

### 2021-05-18 Draft/Draft.Rmd
Update Draft.Rmd

1) over and under represented motifs
```{r}
library(lme4)
d <- read.csv("../Data/Csv/Motifs_frequencies_and_null_models/Motifs_frequency_percentile.csv")
str(d)
head(d)
hist(d$percentil_sizeclass) #wow! 
hist(subset(d, motif == 2)[,"percentil_sizeclass"]) #motif 2 tend to be over-represented
hist(subset(d, motif == 3)[,"percentil_sizeclass"]) #motif 3 tend to be under-represented
#quick and dirty 
mean(subset(d, motif == 3)[,"percentil_sizeclass"]) #mean on 0.18
#more elegant, taking into account the study system
unique(d$Network_id) #we need to recover study_id
m <- lmer(percentil_sizeclass ~ 1+(1|study_id), #HERE WE NEED STUDY SYSTEM ONLY!! 
                       data = subset(d, motif == 2))  
#plot(m)
summary(m) #estimated mean taking into account the random str is 0.XX +- 0.XX (look at the (Intercept) estimate)

#Fig 1 Should be a dot plot (or forest plot) of the mean +- SE of the 15 motifs.
```
2) FG-position associations.
```{r}
library(lme4)
d <- read.csv("../Data/Csv/Motifs_positions_and_null_models/GF_positions_frequency_percentile.csv")
str(d)
head(d)
hist(d$percentil_its_GF) #wow!

#quick and dirty 
mean(subset(d, position == "np3" & Node_FG == 1)[,"percentil_its_GF"], na.rm = TRUE) #mean of 0.61
#more elegant, taking into account the study system
unique(d$Network_id) #we need to recover study_id
m <- lmer(percentil_its_GF ~ 1+(1|study_id), #HERE WE NEED STUDY SYSTEM ONLY!! 
                       data = subset(d, position == "np3" & Node_FG == 1))  
#plot(m)
summary(m) #estimated mean taking into account the random str is 0.XX +- 0.XX (look at the (Intercept) estimate)

>>>>

1) over and under represented motifs
```{r}
library(lme4)
library(tidyverse)
source("../Scripts/Scripts_Alfonso/add_study_id.R")
d <- read_csv("../Data/Csv/Motifs_frequencies_and_null_models/Motifs_frequency_percentile.csv")
d <- add_study_id(d)
str(d)
head(d)

ggplot(d %>% filter(motif != 1), aes(percentil_sizeclass))+
  geom_histogram(color="black", fill="white")+
  facet_wrap(~motif)+
  theme_bw()+
  labs(x="Sizeclass percentile")

# Motifs tend to be over- and under-represensented in real networks

# Mean percentile and SE taking into account the study system
motif_codes <- d %>% filter(motif != 1) %>% # We remove links (motif code: 1)
  select(motif) %>% unique() %>% pull()
motif_means <- tibble(motif = motif_codes)
motif_means$mean <- NA
motif_means$SE <- NA

for(i.motif in 1:length(motif_codes)){

  m <- lmer(percentil_sizeclass ~ 1+(1|study_id), #HERE WE NEED STUDY SYSTEM ONLY!! 
            data = subset(d, motif == motif_codes[i.motif]))

  motif_means$mean[i.motif] <- fixed.effects(m) %>% unname()
  motif_means$SE[i.motif] <- sqrt(diag(vcov(m)))

}

ggplot(NULL) + 
  geom_point(data = d %>% filter(motif != 1),
             aes(x=as.factor(motif), y=percentil_sizeclass, 
                 color = as.factor(motif)),
             position = "jitter",alpha=0.5)+
  geom_errorbar(data = motif_means,aes(x = as.factor(motif), ymin=mean-SE, ymax=mean+SE), 
                width=1.0,size=1)+
  geom_point(data = motif_means,aes(x = as.factor(motif), y=mean), 
                size=2)+
  labs(x="Motif", y = "Percentile")+
  theme_bw()+
  theme(legend.position = "none")

#Fig 1 Should be a dot plot (or forest plot) of the mean +- SE of the 15 motifs.
```
2) FG-position associations.
```{r}
library(lme4)
library(tidyverse)
source("../Scripts/Scripts_Alfonso/add_study_id.R")

all_position_percentiles <- read_csv("../Data/Csv/Motifs_positions_and_null_models/GF_positions_frequency_percentile.csv")

all_position_percentiles <- add_study_id(all_position_percentiles)
str(all_position_percentiles)
head(all_position_percentiles)

list_plants_FG <- as.character(1:10)

plant_position_percentiles_filtered <- all_position_percentiles %>% 
  filter(Node_FG %in% list_plants_FG, !position %in% c("np1","np2"),
         !is.na(expected_freq_its_GF))
pollinator_position_percentiles_filtered <- all_position_percentiles %>% 
  filter(!Node_FG %in% list_plants_FG, !position %in% c("np1","np2"),
         !is.na(expected_freq_its_GF))

ggplot(plant_position_percentiles_filtered, aes(percentil_its_GF))+
  geom_histogram(color="black", fill="white")+
  facet_wrap(~position)+
  theme_bw()+
  labs(x="Sizeclass percentile", title = "Histograms: Percentiles for plant positions")

ggplot(pollinator_position_percentiles_filtered, aes(percentil_its_GF))+
  geom_histogram(color="black", fill="white")+
  facet_wrap(~position)+
  theme_bw()+
  labs(x="Sizeclass percentile",title = "Histograms: Percentiles for pollinator positions")

# estimated mean taking into account the random str is 0.XX +- 0.XX (look at the (Intercept) estimate)

# Plants-------------

# Mean percentile and SE taking into account the study system
plant_codes <- plant_position_percentiles_filtered$position %>% unique()
plant_means <- tibble(position = plant_codes)
plant_means$mean <- NA
plant_means$SE <- NA

for(i.pos in 1:length(plant_codes)){

  m <- lmer(percentil_its_GF ~ 1+(1|study_id), #HERE WE NEED STUDY SYSTEM ONLY!! 
            data = subset(plant_position_percentiles_filtered,
                          position == plant_codes[i.pos]))

  plant_means$mean[i.pos] <- fixed.effects(m) %>% unname()
  plant_means$SE[i.pos] <- sqrt(diag(vcov(m)))

}

ggplot(NULL) + 
  geom_point(data = plant_position_percentiles_filtered,
             aes(x=as.factor(position), y=percentil_its_GF, 
                 color = as.factor(position)),
             position = "jitter",alpha=0.5)+
  facet_wrap(~Node_FG)+
  geom_errorbar(data = plant_means,aes(x = as.factor(position), ymin=mean-SE, ymax=mean+SE), 
                width=1.0,size=1)+
  geom_point(data = plant_means,aes(x = as.factor(position), y=mean), 
                size=2)+
  labs(x="Position", y = "Percentile",title = "Pollinators")+
  theme_bw()+
  theme(legend.position = "none",axis.text.x=element_text(angle=90,vjust=0.5, hjust=1))

# Pollinators--------

# Mean percentile and SE taking into account the study system
pollinator_codes <- pollinator_position_percentiles_filtered$position %>% unique()
pollinator_means <- tibble(position = pollinator_codes)
pollinator_means$mean <- NA
pollinator_means$SE <- NA

for(i.pos in 1:length(pollinator_codes)){

  m <- lmer(percentil_its_GF ~ 1+(1|study_id), #HERE WE NEED STUDY SYSTEM ONLY!! 
            data = subset(pollinator_position_percentiles_filtered,
                          position == pollinator_codes[i.pos]))

  pollinator_means$mean[i.pos] <- fixed.effects(m) %>% unname()
  pollinator_means$SE[i.pos] <- sqrt(diag(vcov(m)))

}

ggplot(NULL) + 
  geom_point(data = pollinator_position_percentiles_filtered,
             aes(x=as.factor(position), y=percentil_its_GF, 
                 color = as.factor(position)),
             position = "jitter",alpha=0.5)+
  facet_wrap(~Node_FG)+
  geom_errorbar(data = pollinator_means,aes(x = as.factor(position), ymin=mean-SE, ymax=mean+SE), 
                width=1.0,size=1)+
  geom_point(data = pollinator_means,aes(x = as.factor(position), y=mean), 
                size=2)+
  labs(x="Position", y = "Percentile", title = "Pollinators")+
  theme_bw()+
  theme(legend.position = "none",axis.text.x=element_text(angle=90,vjust=0.5, hjust=1))
